|

### 2023-01-27 paper.Rmd
Fri Jan 27 16:00:52 CET 2023: update-remote.sh

Figure \@ref(fig:airfoil_median_fitness_30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset. The results are aggregated for $10$ individuals runs per algorithm and also include regular GP as a benchmark. For all three algorithms an improvement in the best found fitness can be observed during the evolutionary search but the results show that DAE-GP based algorithms, on average, find solutions with much higher fitness as traditional GP.

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better performance than regular DAE-GP for both the training and the testing set in overall fitness. Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.

```{r airfoil_median_fitness_30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil", out.width="90%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
```

![Best Fitness Distribution after 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png)
lalal

![Population Diversity over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens.png)
Since @dae-gp_2022_symreg shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.

![Average Solution Size over 30 Generations - Airfoil](./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png)

>>>>

### Airfoil Dateset - 2 hidden Layers

Figure \@ref(fig:airfoil-median-fitness-30gens) shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks. The results are aggregated for $10$ individuals runs per algorithm and also include regular GP as a benchmark. For all three algorithms an improvement in the best found fitness can be observed during the evolutionary search but the results show that DAE-GP based algorithms, on average, find solutions with much higher fitness as traditional GP.

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness. Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.

```{r airfoil-median-fitness-30gens, echo=FALSE, fig.cap="Best Fitness over 30 Generations - Airfoil (2HL)", out.width="90%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png")
```
The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure  \@ref(fig:airfoil-distribution-fitness-30gens). The distribution for pre-trained DAE-GP shows to be preferable to that of regular DAE-GP on both sets since it shows a lower median (med.) as well as less dispersion measured by the standard deviation (std.).

```{r airfoil-distribution-fitness-30gens, echo=FALSE, fig.cap="Best Fitness after 30 Generations - Airfoil (2HL)", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png")
```

Since previous work by @dae-gp_2022_symreg demonstrated that DAE-GP was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.

Since @dae-gp_2022_symreg shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP. Figure \@ref(fig:airfoil-size-combined-30gens) shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population. It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.

```{r airfoil-size-combined-30gens, echo=FALSE, fig.cap="Median Solution Size over 30 Generations - Airfoil (2HL)", out.width="80%"}
knitr::include_graphics("./img/airfoil_2hl_maxIndSize_fullRun_30gens/size_combined_median.png")
```

### 2023-01-27 paper.tex
Fri Jan 27 16:00:52 CET 2023: update-remote.sh

After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP. Again the Airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]}, the parameters are described in table

@@

\begin{figure}
\centering
\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens.png}
\caption{Best Fitness over 30 Generations - Airfoil}
\end{figure}

\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot.png}
lalal

\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_popDiversity_byGens.png}
Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP.

\begin{figure}
\centering
\includegraphics{./img/airfoil_2hl_maxIndSize_fullRun_30gens/mean_Size_byGens.png}
\caption{Average Solution Size over 30 Generations - Airfoil}
\end{figure}

>>>>

After closely examining the effect of pre-training on DAE-GPs generalization behaviour another series of experiments is conducted to study how pre-training influences the overall search behaviour of DAE-GP. The airfoil problem was selected as a first benchmark problem to test pre-trained DAE-GP against traditional DAE-GP.

@@

\hypertarget{airfoil-dateset---2-hidden-layers}{%
\subsubsection{Airfoil Dateset - 2 hidden Layers}\label{airfoil-dateset---2-hidden-layers}}

Figure \ref{fig:airfoil-median-fitness-30gens} shows both the mean and median of the best fitness for each generation on both the test and training set of the airfoil dataset for using 2 hidden layers inside the DAE-LSTM networks. The results are aggregated for \(10\) individuals runs per algorithm and also include regular GP as a benchmark. For all three algorithms an improvement in the best found fitness can be observed during the evolutionary search but the results show that DAE-GP based algorithms, on average, find solutions with much higher fitness as traditional GP.

The first interesting observation from this experiment is, that the pre-trained DAE-GP shows better mean and median performance than regular DAE-GP for both the training and the testing set in overall fitness. Regarding the presence of overfitting, both DAE-GP variants show only a very small gap between the testing and training fitness which indicates a good generalization behaviour.

{\centering \includegraphics[width=0.9\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens_withGP/mean_median_fitness_byGens}

\caption{Best Fitness over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-median-fitness-30gens}
\end{figure}

The distribution of final fitness scores for pre-trained and regular DAE-GP are visualized as boxplots in figure \ref{fig:airfoil-distribution-fitness-30gens}. The distribution for pre-trained DAE-GP shows to be preferable to that of regular DAE-GP on both sets since it shows a lower median (med.) as well as less dispersion measured by the standard deviation (std.).

{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens/final_fit_boxplot}

\caption{Best Fitness after 30 Generations - Airfoil (2HL)}\label{fig:airfoil-distribution-fitness-30gens}
\end{figure}

Since previous work by {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} demonstrated that DAE-GP was capable of producing solutions that are much smaller in tree size that regular GP, I also studied the influence of pre-training on the size of solutions found by DAE-GP.

Since {[}\protect\hyperlink{ref-dae-gp_2022_symreg}{17}{]} shows, that DAE-GP finds higher quality that are also smaller in size than those found by average GP for a given number of 10.000 fitness evaluations, I also studied the differences in average solution size between pre-trained and regular DAE-GP. Figure \ref{fig:airfoil-size-combined-30gens} shows the median of both the average size of individuals inside the population as well as the size of the currently best performing solution inside the population. It can be observed that both DAE-GP variants strongly reduce the average solutions size already in the first generations and that the average solution size inside the population constantly stays at this very low level for the rest of the search.

{\centering \includegraphics[width=0.8\linewidth]{./img/airfoil_2hl_maxIndSize_fullRun_30gens/size_combined_median}

\caption{Median Solution Size over 30 Generations - Airfoil (2HL)}\label{fig:airfoil-size-combined-30gens}
\end{figure}
